---
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../"
    )
  })
title: "Wine Dataset EDA"
author: "Sultan Alkadhi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# wine <- read_csv(here('data/winequality-red-kaggle.csv'))
```



# Data 
Include a static markdown table describing the name of each variable and what it is.
#EDA
Include appropriate data visualizations from your EDA and text explaining what you observed.
#Conclusion
Include a short text describing your overall insights drawn from examining the data.


When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
